---
title: "Non-convex trender filter using proximal algorithms"
author: "Lei Sun"
date: "2019-04-08"
output:
  workflowr::wflow_html:
    code_folding: hide
---

## Introduction

The objective is as follows
$$
\min\limits_{\beta} \frac12\|y - X\beta\|_2^2 + \lambda\Phi(D\beta)
$$
where $\Phi(\cdot): \mathbb{R}^m \to \mathbb{R}$ is a penalty function in the form of
$$
\Phi(x) := \sum_{i = 1}^m\phi(x_i)
$$
where $\phi(\cdot): \mathbb{R} \to \mathbb{R}$ can be in many forms. Note that when $\phi(\cdot)$ is convex, the whole objective is convex. The two representative examples considered here are $L_1$: $\phi(\cdot)=|\cdot|$ (convex) and double-pareto: $\phi(\cdot) = \log(1+|\cdot|)$ (nonconvex). The $L_1$ case has been solved by [`genlasso`](https://github.com/glmgen/genlasso).

## Algorithm

## Initial simulation
$X, D, y$ consist of random numbers. $\lambda = 10$.

```{r}
source("../code/bayesprox.R")
```

```{r}
set.seed(7)

X <- matrix(rnorm(100 * 80), 100)
D <- matrix(rnorm(50 * 80), 50)
y <- rnorm(100)

fit.bayesprox <- bayesprox(y, X, D, lambda = 10, penalty = "l1")
fit.genlasso <- genlasso::genlasso(y, X, D)
beta.bayesprox <- fit.bayesprox$beta
beta.genlasso <- coef(fit.genlasso, lambda = 10)$beta
plot(beta.bayesprox, beta.genlasso,
     xlab = expression(paste(hat(beta), " by bayesprox")),
     ylab = expression(paste(hat(beta), " by genlasso")),
     main = expression(paste(L[1], "-penalty")))
abline(0, 1)
```

## Piecewise constant trend filter
$X = I, D = $ 1-difference matrix. $\lambda$ is chosen by `genlasso` using CV.

```{r}
set.seed(7)
n <- 1e2
p <- 0.95
sigma <- 0.5
b <- 5

y <- x <- v <- c()
x[1] <- runif(1, -b, b)
for (i in 1 : (n - 1)) {
  v[i] <- rbinom(1, 1, 1 - p) * runif(1, -b, b)
  x[i + 1] <- x[i] + v[i]
}
y <- x[1 : n] + rnorm(n, 0, sigma)
plot(y, pch = 19, cex = 0.5, col = "grey50", ylab = "")
points(x, pch = 19, cex = 0.5, type = "l")
legend("bottomleft", lty = 1, "Truth")

X <- diag(n)
D <- genlasso::getD1d(n)

fit.genlasso = genlasso::trendfilter(y, ord = 0)
cv = genlasso::cv.trendfilter(fit.genlasso)
lambda <- cv$lambda.1se
beta.genlasso <- coef(fit.genlasso, lambda)$beta
```

```{r}
fit.bayesprox <- bayesprox(y, X, D, lambda, penalty = "dp", par = 1)
beta.bayesprox <- if (fit.bayesprox$converged) fit.bayesprox$beta else fit.bayesprox$beta.mat[, which.min(fit.bayesprox$obj_vec)]
# beta.bayesprox <- fit.bayesprox$beta

plot(y, pch = 19, cex = 0.5, col = "grey50", ylab = "")
points(x, pch = 19, cex = 0.5, type = "l")
points(beta.genlasso, pch = 19, cex = 0.5, col = "green", type = "l")
points(beta.bayesprox, pch = 1, cex = 0.5, col = "blue", type = "l")
legend("bottomleft", lty = 1, c("Truth", "L1 by genlasso", "dp by bayesprox"), col = c("black", "green", "blue"))
```

```{r}
cat("MSE of y")
sqrt(mean((y - x)^2))
cat("MSE of genlasso")
sqrt(mean((beta.genlasso - x)^2))
cat("MSE of bayesprox")
sqrt(mean((beta.bayesprox - x)^2))

cat("Objective of y")
obj_fun(y, y, X, D, lambda, penalty = "dp", par = 1)
cat("Objective of truth")
obj_fun(x, y, X, D, lambda, penalty = "dp", par = 1)
cat("Objective of genlasso")
obj_fun(beta.genlasso, y, X, D, lambda, penalty = "dp", par = 1)
cat("Objective of bayesprox")
obj_fun(beta.bayesprox, y, X, D, lambda, penalty = "dp", par = 1)
```

## Piecewise linear trend filter
$X = I, D = $ 2-difference matrix. $\lambda$ is chosen by `genlasso` using CV.

```{r}
set.seed(7)
n <- 1e2
p <- 0.95
sigma <- 1.5
b <- 0.5

y <- x <- v <- c()
x[1] <- 0
v[1] <- runif(1, -b, b)
for (i in 1 : (n - 1)) {
  x[i + 1] <- x[i] + v[i]
  ind <- rbinom(1, 1, p)
  v[i + 1] <- ind * v[i] + (1 - ind) * runif(1, -b, b)
}
y <- x[1 : n] + rnorm(n, 0, sigma)
plot(y, pch = 19, cex = 0.5, col = "grey50", ylab = "")
points(x, pch = 19, cex = 0.5, type = "l")
legend("topright", lty = 1, "Truth")

X <- diag(n)
D <- genlasso::getDtf(n, 1)

fit.genlasso = genlasso::trendfilter(y, ord = 1)
cv = genlasso::cv.trendfilter(fit.genlasso)
lambda <- cv$lambda.1se
beta.genlasso <- coef(fit.genlasso, lambda)$beta
```

```{r}
fit.bayesprox <- bayesprox(y, X, D, lambda, penalty = "dp", par = 1)
beta.bayesprox <- if (fit.bayesprox$converged) fit.bayesprox$beta else fit.bayesprox$beta.mat[, which.min(fit.bayesprox$obj_vec)]
# beta.bayesprox <- fit.bayesprox$beta

plot(y, pch = 19, cex = 0.5, col = "grey50", ylab = "")
points(x, pch = 19, cex = 0.5, type = "l")
points(beta.genlasso, pch = 19, cex = 0.5, col = "green", type = "l")
points(beta.bayesprox, pch = 1, cex = 0.5, col = "blue", type = "l")
legend("topright", lty = 1, c("Truth", "L1 by genlasso", "dp by bayesprox"), col = c("black", "green", "blue"))
```

```{r}
cat("MSE of y")
sqrt(mean((y - x)^2))
cat("MSE of genlasso")
sqrt(mean((beta.genlasso - x)^2))
cat("MSE of bayesprox")
sqrt(mean((beta.bayesprox - x)^2))

cat("Objective of y")
obj_fun(y, y, X, D, lambda, penalty = "dp", par = 1)
cat("Objective of truth")
obj_fun(x, y, X, D, lambda, penalty = "dp", par = 1)
cat("Objective of genlasso")
obj_fun(beta.genlasso, y, X, D, lambda, penalty = "dp", par = 1)
cat("Objective of bayesprox")
obj_fun(beta.bayesprox, y, X, D, lambda, penalty = "dp", par = 1)
```
